Practical ML - Coursera
========================================================

## Loading libs and reading data
```{r echo=FALSE}
library(caret)
pml.training <- read.csv("~/2014/coursera/practical_ml/pml-training.csv",na.strings = c("NA",""))
pml.testing <- read.csv("~/2014/coursera/practical_ml/pml-testing.csv",na.strings = c("NA",""))
```


## Setting seed 
```{r}
set.seed(123123)
```

## Creating training/test partitions
```{r}
inTrain <- createDataPartition(y=pml.training$classe, p=0.7, list=FALSE)
training <- pml.training[inTrain,]
testing <- pml.training[-inTrain,]
```

## Cleaning Data
* Remove columns related to user name, timestamps and window (columns 1:7)
* Check and remove columns with almost zero var
* Remove columns with almost all missing values
```{r}
n_feats = dim(pml.training)[2]
names(training[,1:7])
training = training[,8:n_feats] 
testing = testing[,8:n_feats]
pml.testing = pml.testing[,8:n_feats] 

n_train = dim(training)[1]
nzv = nearZeroVar(training)
dim(training)
training = training[,-nzv]
testing = testing[,-nzv]
dim(training)
nas = colSums(is.na(training))/n_train
table(nas)
training = training[,names(nas[nas==0])]
testing = testing[,names(nas[nas==0])]
dim(training)

pml.testing = pml.testing[,-nzv]
pml.testing = pml.testing[,names(nas[nas==0])[1:52]]
```

### Feature selection with PCA
```{r}
pca = preProcess(x=training[,-which(names(training) == "classe")],method='pca',thresh=.9)
pca
trainPC <- predict(pca,training[,-which(names(training) == "classe")])
```

### Model with cross-validation
* Check in-sample error
* Check cv error
```{r}
fitControl <- trainControl(method="repeatedcv", number=10, repeats=10)
model <- train(training$classe ~ ., data=trainPC, method="rf", ntree=25, trControl=fitControl)
confusionMatrix(training$classe,predict(model,trainPC)) # in-sample error
model # cv info and error

```

### Testing
```{r}
testPC <- predict(pca,testing[,-which(names(training) == "classe")])
confusionMatrix(testing$classe,predict(model,testPC))
```

### Generating predictions
```{r}
pml.testingPC <- predict(pca,pml.testing)
ans = predict(model,pml.testingPC)
ans

pml.testPC <- predict(pca,pml.testing)
ans = predict(model,pml.testPC)

pml_write_files = function(x) {
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}

pml_write_files(ans)

```